Applying Pose Estimation to Predict the Outcome of A Free Throw

### 2022-12-01 Papers/Report/report.tex
Finish first draft of paper

%%%%%%%%% TITLE - PLEASE UPDATE
\title{Applying Pose Estimation to Predict the Outcome of Basketball Shots}

@@

\section{Analysis}
From our experiments we concluded that our second approach to feature generation and classification was
far more successful than our first approach. We believe that the main issue with our first approach was
insufficient data. We were only able to generate about 300 total clips, which meant we had about 210
feature vectors to train on and 90 to test on. This simply was not enough data to train a robust classifier.

@@

\section{Conclusion}

>>>>

%%%%%%%%% TITLE - PLEASE UPDATE
\title{Applying Pose Estimation to Predict the Outcome of A Free Throw}

@@

\subsection{Other Explorations}
After achieving these results, we explored other methods that have historically been useful in improving the performance of
classification models. Principal Component Analysis (PCA) was the first method we explored to attempt to find the most significant
features of our features. However, when we performed the PCA transform, we greatly hindered the performance of our model to about the
performance of random guessing.

Additionally, we attempted stacking models together to improve our performance. We stacked many different classification models together
and attempted to use the majority voting method to improve the performance. However, this also hindered the performance of our model.

\section{Analysis}
\subsection{Performance Analysis}
From our experiments we concluded that our second approach to feature generation and classification was
far more successful than our first approach. We believe that the main issue with our first approach was
insufficient data. We were only able to generate about 300 total clips, which meant we had about 210
feature vectors to train on and 90 to test on. This simply was not enough data to train a robust classifier.

@@

\subsection{Pitfalls and Issues}
Although we were able to achieve adequate performance metrics with our approach, there are still some issues that
our current method cannot overcome.

\subsubsection{Invariance to Camera Position}

One of the main issues with our current approach is that our feature vector generation highly sensitive to the position of our camera. For our data collection,
we held the camera in a fixed position to record our test subject shooting free throws, and our feature vectors used the absolute coordinates of each of the 17 key points
to create our feature vector for a frame. This means that if we were to move the camera, or even if the person were to move, the feature vectors would be completely different, 
even if the shooting form is the same.

In order to fix this issue, we could use relative coordinates, fixing one of the key points as the origin, and calculating the relative coordinates of the other 16 key points.
This would allow our feature representation to be more invariant to the position of the camera or the test subject. Additionally, we could use a feature representation that took into
account the angles between the key points because the angles of joints are very important in determining the form of a shot. Using the angles would also make the feature representation
not depend on the absolute position.

\subsubsection{Loss of Temporal Information}

In our current representation, we are performing a majority vote on all the frames of a clip, but when we do this, we lose all temporal information and the sequential nature of the shot.
This means, theoretically, we could shuffle the frames of the clip, or even reverse the video, and the model will still predict the same.

In order to capture the sequential
nature of the shooting form, it would have been beter to potentially use a recurrent neural network to capture the temporal information.

\section{Conclusion and Future Works}
In this paper, we attempted to leverage pose estimation to help predict the outcome of a free throw. This involved collecting data
of a test subject shooting free throws, and using the pose estimation to generate feature vectors, and training a 
classification model to predict the outcome of a shot.

Through this process, we learned a great deal about how pose estimation works and the methodology behind generating feature vectors from data as well as
the nuances of training classification models.

Future works for this project would be to imrpove the robustness of our feature vector generation and capturing the temporal information, as explained 
in the previous section. Additionally, we could explore extending this project to work at different locations on the court and for different people shooting.
Lastly, we could explore the possibility of performing the prediction in real time, so that the results can be produced live instead of from a video recording.
